Categories of Optics

### 2018-04-18 paper.tex
More diagrams

% Categories
\newcommand{\Set}{\mathbf{Set}}
\newcommand{\Cat}{\mathbf{Cat}}
\newcommand{\Prof}{\mathbf{Prof}}
\newcommand{\MonCat}{\mathbf{MonCat}}
\newcommand{\LaxMonCat}{\mathbf{LaxMonCat}}
\newcommand{\SymmMonCat}{\mathbf{SymmMonCat}}
\newcommand{\Tele}{\mathbf{Tele}}

@@

@@

\todo{
\subsection{(Co)ends and (Co)Yoneda}
}

@@

We begin by constructing the category of optics in a symmetric monoidal category $\C$. This category was first defined in \cite[Section 6]{Doubles} to investigate the relationship between Tambara modules and the center of a monoidal category. There it is called the \emph{double} $\Double_l \C$ of a monoidal category, and our definition is identical, other than flipping the direction of the morphisms to match the existing work on lenses.

@@

@@

\begin{remark}
Note that the definition of the $\Optic$ category involves a coend possibly indexed by a large category. To be maximally careful, we should only ever discuss $\Optic$ categories where we know that the coend exists by some other means, e.g., by exhibiting an isomorphism with some other set. For many---but not all---of the examples given later, we provide such an isomorphism.

@@

\begin{proposition}
$\iota : \C \times \C^\op \to \Optic_\C$ is a strong monoidal functor. 
\end{proposition}
\begin{proposition}
\todo{todo}
\end{proposition}

@@

In particular, for any $M \in \C$, the image of the identity yields an optic $c_M : (M, M) \hto (I, I)$, that we will call the connector.

States are not as easy to describe in general, however:

@@

\begin{proposition}
Together this operation defines a \todo{pseudo?}functor $\Optic : \SymmMonCat \to \SymmMonCat$.
\end{proposition}
\begin{proof}
\todo{todo}
\end{proof}

@@

Categories of optics admit their own diagrammatic language, by virtue of them being \emph{teleological categories}. In fact, we show in Theorem \ref{optic-is-free-teleological-cat} that $\Optic_\C$ is the free teleological category on a collection of dualisable morphisms $\C$.

In this section we assume that $\C$ is a \emph{strict} symmetric monoidal category.

\begin{definition}[{\cite[Definition 5.1]{CoherenceForLenses}}]
A \emph{teleological category is} a symmetric monoidal category $(\C, \otimes, I)$, equipped with:
\begin{itemize}
\item A wide symmetric monoidal subcategory $\C_d$ of \emph{dualisable morphisms}, with an involutive strong symmetric monoidal functor $(-)^* : \C_d \to \C_d^\op$; and,
\item A family of morphisms $\varepsilon_X : X \otimes X^* \to I$, called \emph{counits}, natural with respect to the morphisms in $\C_d$, such that
\[
\begin{tikzcd}
X \otimes Y^* \ar[r, "f \otimes Y^*"]  \ar[d, "X \otimes f^*", swap] & Y \otimes Y^* \ar[d, "\varepsilon_Y"] \\
X \otimes X^* \ar[r, "\varepsilon_X", swap] & I
\end{tikzcd} \hspace{1cm}
\begin{tikzcd}
X^* \otimes X \ar[r, "s_{X^*, X}"]  \ar[dr, "\varepsilon_{X^*}", swap] & X \otimes X^* \ar[d, "\varepsilon_X"] \\
& I
\end{tikzcd} \hspace{1cm}
\begin{tikzcd}
X \otimes Y \otimes X^* \otimes Y^* \ar[r, "X \otimes s_{Y,X^*} \otimes X^*"]  \ar[dr, "\varepsilon_{X \otimes Y}", swap] & X \otimes X^* \otimes Y \otimes Y^* \ar[d, "\varepsilon_X \otimes \varepsilon_Y"] \\
& I
\end{tikzcd}
\]
commute for all $f : X \to Y$ is in $\C_d$.
\end{itemize}
\end{definition}
Note that because $\C_d$ is symmetric monoidal and has the same collection of objects as $\C$, the symmetric monoidal structure of $\C$ must all be contained in $\C_d$.

\begin{example}
Any symmetric monoidal category with terminal monoidal unit is trivially teleological, setting the dualisable morphisms to be only the structure maps of the monoidal category.
\end{example}

@@

\begin{remark}
Compact closed categories very nearly form an example of teleological categories. In a compact closed category, there are canonical isomorphisms $(X \otimes Y)^* \cong Y^* \otimes X^*$. In a teleological category, however, duality does not reverse the order of the tensor. \todo{If $\C$ is strict though, does this matter?}
\end{remark}

@@

Together we have $\Tele$, the (1-)category of teleological categories and functors. There are evident functors $U : \Tele \to \SymmMonCat$ and $(-)_d : \Tele \to \SymmMonCat$, that take a teleological category to its underlying symmetric monoidal category and subcategory of dualisable morphisms respectively.

\begin{proposition}
$\Optic_\C$ forms a teleological category, where:
\begin{itemize}
\item The dualisable morphisms are all morphisms of the form $\iota(f, g)$;
\item The involution is given on objects by $(S, S')^* = (S', S)$, and on a morphism $\iota(f, g)$ by $\iota(g, f)$;
\item The counit $\varepsilon_{(S, S')}$  is given by the connector $c_{S \otimes S'}$.
\end{itemize}
\end{proposition}
\begin{proof}
That morphisms of the form $\iota(f, g)$ constitute a category and that $(-)^*$ is a symmetric monoidal involution is clear. \todo{actually check monoidal part}

To check extranaturality, suppose we have a dualisable optic $\iota(f, g) : (X, X') \to (Y, Y')$, so $f : X \to Y$ and $g : Y' \to X'$. Extranaturality is witnessed by the equality of the string diagrams:
\begin{center}
\includegraphics[width=.4\textwidth]{diagrams/counit-extranaturality}
\end{center}
Symmetry:
\begin{center}
\includegraphics[width=.4\textwidth]{diagrams/counit-symmetry}
\end{center}
Monoidal:
\begin{center}
\includegraphics[width=.4\textwidth]{diagrams/counit-monoidal}
\end{center}

@@

\todo{They stress about the empty set, have I avoided that issue?}

\begin{proposition}
$\Optic : \SymmMonCat \to \Tele$ is a functor.
\end{proposition}
\begin{proof}
We have to show that for a symmetric monoidal functor $F : \C \to \D$, the induced functor $\Optic_\C \to \Optic_\D$ is teleological.

@@

\begin{proposition}
\label{prop-optic-decompose}
Suppose $p : (S, S') \to (A, A')$ has representative $\langle l, r \rangle$. Then 
\begin{align*}
p = (\varepsilon_{(M, I)} \otimes (A, A'))\iota(l, r).
\end{align*}
\end{proposition}
\begin{proof}
\todo{by string diagram}
\end{proof}

\begin{theorem}
\label{optic-is-free-teleological-cat}
Suppose $\C$ is a \todo{strict} symmetric monoidal category and $\T$ is a teleological category. For all monoidal functors $F : \C \to \T_d$, there exists a teleological functor $K : \Lens_\C \to \T$, unique up to isomorphism, such that $Kj \cong F$.
%$\Optic$ is left adjoint to $(-)_d$, so $\Optic_\C$ is the free teleological category on a category of dualisable morphisms $\C$.
\end{theorem}
\begin{proof}
\todo{This is the point at which I have run in to trouble: TODO}

%On objects, define $K(S, S') = FS \otimes (FS')*$. Suppose $p : (S, S') \hto (A, A')$ is an optic with representative $\langle l, r \rangle$. By Proposition \ref{prop-optic-decompose} and the previous Lemma, $p$ may be decomposed as:
%\begin{align*}
%\langle l, r \rangle = (\varepsilon_{(M, I)}\iota(\rho_M^{-1}, \lambda_M) \otimes \id_{(A, A')})\iota(l, s_{M,A'}r)
%\end{align*}
%But now, in order for $K$ to be a teleological functor with $Kj \cong F$, each piece of this expression must be preserved by $K$, so we define:
%\begin{align*}
%K\langle l, r \rangle = (\varepsilon_{FM} \otimes \id_{FA \otimes FA'})(Fl \otimes s_{FM,FA'} (Fr)^* )
%\end{align*}
%
%This is much easier to understand in the diagrammatic language for teleological categories: \todo{todo}
%
%We now have to check that this is well defined. \todo{todo}
%
%\todo{incomplete}
\end{proof}

@@

\begin{itemize}
\item Given $f : S \to A$ and $g : A' \to S'$, the optic $j(f) \otimes j(g)^* : (S \otimes I, I \otimes S') \hto (A \otimes I, I \otimes A')$ is not equal to $\iota(f, g)$. It is also not in a form that can be composed with the unitors of $\Optic_\C$, as the $I$s are not all on the same side.
\item \todo{wasn't there more?}
\end{itemize}

@@

For a fixed $M \in \M$, there is a map $\C(S, M A) \times \C(M A, S) \to \C(S, S)$ given by composition. This induces a map $outside : \Optic_\M(S, A) \to \C(S, S)$. By composition in the other direction we have a map $\C(S, M A) \times \C(M A, S) \to \C(M A, M A)$. Including this object $\C(M A, M A)$ into the coend $\int^{M \in \M} \C(M A, M A)$, we induce a map $inside : \Optic_\M(S, A) \to \int^{M \in \M} \C(M A, M A)$. 
\begin{definition}
An optic $p : S \hto A$ is \emph{lawful} if $outside(p) = \id_S$, and $inside(p) = \langle t A \rangle$ for some $t : M \to M$ in $\M$.
\end{definition}

@@

\begin{proposition}
\label{prop-onthenose}
If $p : S \to A$ is a lawful optic, then there exists a representative $p = \langle l, r \rangle$ such that $rl = \phi A$ on the nose for some $\phi : M \to M$.
\end{proposition}
\begin{proof}
\todo{This is painful but seems important. There may be a more general principle at work here but I couldn't see it. Put in an appendix?}

There is a well known formula for the coend as the coequaliser in the diagram
\[
\begin{tikzcd}
\coprod_{M \to N} P(N, M) \ar[r,shift left=.75ex]  \ar[r,shift right=.75ex] & \coprod_{M \in \M} P(M, M) \ar[r] & \int^{M \in \M} P(M, M)
\end{tikzcd}
\]
where $P$ is a functor $\M^\op \times \M \to \E$.

In our case $\E = \Set$, and unwinding the coequaliser this states that $\int^{M \in \M} \C(M A, M A)$ is a quotient of the set $\coprod_{M \in \M} \C(M A, M A)$. Two morphisms $f : M A \to M A$ and $g : N A \to N A$ are identified when there exists a $k : N A \to M A$ and natural transformation $\phi : M \to N$ such that the diagram
\[
\begin{tikzcd}
M A \ar[r, "\phi_A"] \ar[d, "f", swap] & N A \ar[d, "g"] \ar[dl, "k"] \\
M A \ar[r, "\phi_A", swap] & N A
\end{tikzcd}
\]
commutes. This gives a binary relation $f \rightsquigarrow g$ that is not likely to be symmetric or transitive. The coend is obtained by identifying all morphisms related in this way, so in all, $\langle f \rangle = \langle g \rangle$ iff there exists a zigzag $f \leftrightsquigarrow u_1 \leftrightsquigarrow \dots \leftrightsquigarrow u_n \leftrightsquigarrow g$ where the relation may apply in either direction. \todo{replace this with $f = u_1 \leftrightsquigarrow \dots u_n = g$ for clarity}

If $f \rightsquigarrow g$ then $f^2 \rightsquigarrow g^2$: this is clear by stacking the above diagram, and then taking $k' = fk = kg$. This argument extends to show that $f^n \rightsquigarrow g^n$ for any $n$.

Similarly unwinding the coequaliser for $\int^{M \in \M} \C(S, M A) \times \C(M A, S)$, two pairs $(l, r) : (S \to M A, MA \to S)$ and $(l', r') : (S \to NA, NA \to S)$ in $\coprod_{M \in \M} \C(S, M A) \times \C(M A, S)$ are identified when there exists a $\phi : M \to N$ such that
\[
\begin{tikzcd}
M A \ar[r, "\phi_A"] \ar[d, "r", swap] & N A \ar[d, "r'"] \\
S \ar[r, equal] \ar[d, "l", swap] & S \ar[d, "l'"]  \\
MA \ar[r, "\phi_A"] & N A
\end{tikzcd}
\]
commutes. \todo{put the $S$ on the outside?}

We will now prove the result by induction. Because $inside(p) = \langle t \rangle$, there is a representative $\langle l, r \rangle$ so that $\langle lr \rangle = \langle t \rangle$. There therefore must exist a finite chain of relations $lr \leftrightsquigarrow u_1 \leftrightsquigarrow \dots \leftrightsquigarrow u_n \leftrightsquigarrow t$.
Supposing that the first relation faces rightwards, we have a diagram:
\[
\begin{tikzcd}
M A \ar[r, "\phi_A"] \ar[d, "r", swap] & N A \ar[dd, "u_1"] \ar[ddl, "k"] \\
S \ar[d,"l",swap] & \\
M A \ar[r, "\phi_A", swap] & N A
\end{tikzcd}
\]
Note that $r = rlr = rk\phi_A$. We therefore have the commutative diagram
\[
\begin{tikzcd}
M A \ar[r, "\phi_A"] \ar[d, "r", swap] & N A \ar[d, "rk"] \\
S \ar[r, equal] \ar[d, "l", swap] & S \ar[d, "\phi_Al"] \\
MA \ar[r, "\phi_A", swap] & N A
\end{tikzcd}
\]
where the composite of the morphisms on the right is $\phi_A l r k = \phi_A k \phi_A k = u_1^2$. Composing in the other direction we have $r k \phi_A l = rlrl = \id_S$. We have shown that whenever $rl = \id_S$ and $lr \rightsquigarrow u_1$, then there exist $l_1$ and $r_1$ so that $r_1l_1 = \id_S$, $l_1r_1 = u_1^2$ and $(l, r) \rightsquigarrow (l_1, r_1)$. A symmetric argument shows that if instead $lr \leftsquigarrow g$, there exist $l_1$ and $r_1$ so that $r_1l_1 = \id_S$, $l_1r_1 = u_1^2$ and $(l, r) \leftsquigarrow (l_1, r_1)$.

We can now inductively apply the above argument to the chain $u_1^2 \leftrightsquigarrow \dots \leftrightsquigarrow u_n^2 \leftrightsquigarrow t^2 = t$, obtained by squaring each morphism in the original chain. We eventually obtain a chain $(l, r) \leftrightsquigarrow (l_1, r_1)\leftrightsquigarrow \dots \leftrightsquigarrow (l_n, r_n) \leftrightsquigarrow (l_t, r_t)$, where $r_tl_t = \id_S$ and $l_t r_t = t$. This pair $(l_t, r_t)$ is the required representative.
\end{proof}

@@

@@

\begin{align*}
\Optic_\otimes((S, S'), (A, A')) &= \int^{M \in \C} \C(S, M \otimes A) \times \C(M \otimes A', S') \\
&\cong \int^{M \in \C} \C(\homC(S,A), M) \times \C(M \otimes A', S') \\
&\cong \C(\homC(S, A) \otimes A', S')
\end{align*}
where $\homC(S, A)$ denotes the internal hom.

@@

>>>>

% Categories
\newcommand{\Set}{\mathbf{Set}}
\newcommand{\Cat}{\mathbf{Cat}}
\newcommand{\Prof}{\mathbf{Prof}}
\newcommand{\Core}{\mathbf{Core}}
\newcommand{\MonCat}{\mathbf{MonCat}}
\newcommand{\LaxMonCat}{\mathbf{LaxMonCat}}
\newcommand{\SymmMonCat}{\mathbf{SymmMonCat}}
\newcommand{\Tele}{\mathbf{Tele}}

@@

\newcommand{\switched}{\mathbin{\tilde{\otimes}}}

@@

\todo{
\subsection{(Co)ends and (Co)Yoneda}
}
If $\E$ is cocomplete, there is a formula for the coend as the coequaliser in the diagram
\[
\begin{tikzcd}
\coprod_{M \to N} P(N, M) \ar[r,shift left=.75ex]  \ar[r,shift right=.75ex] & \coprod_{M \in \M} P(M, M) \ar[r] & \int^{M \in \M} P(M, M)
\end{tikzcd}
\]
where $P$ is a functor $\M^\op \times \M \to \E$.

@@

We begin by constructing the category of optics in a symmetric monoidal category $\C$. This category was first defined in \cite[Section 6]{Doubles} to investigate the relationship between Tambara modules and the center of a monoidal category. There it is called the \emph{double} $\Double_l \C$ of a monoidal category, and our definition is identical, other than flipping the direction of the morphisms to match the existing work on lenses, and restricting our attention to the unenriched case.

@@

\begin{example}
For any three objects $M, A, A' \in \C$, there is the \emph{tautological} optic \[t_{M,A} : (M \otimes A, M \otimes A') \hto (A, A')\] which is given by $\langle \id_{M \otimes A}, \id_{M \otimes A'} \rangle$.
\end{example}

@@

\begin{remark}
Note that the definition of the $\Optic$ category involves a coend possibly indexed by a large category. To be careful, we should only ever discuss $\Optic$ categories where we know that the coend exists by some other means, e.g., by exhibiting an isomorphism with some other set. For many---but not all---of the examples given later, we provide such an isomorphism.

@@

\begin{proposition}
$\iota : \C \times \C^\op \to \Optic_\C$ is a strong monoidal functor. 
\end{proposition}
\begin{proof}
\todo{todo}
\end{proof}

@@

In particular, for any $S \in \C$, the image of the identity yields an optic $c_S : (S, S) \hto (I, I)$, that we will call the connector:
\begin{center}
\input{diagrams/connector-full.tikz}
\end{center}
Or, again omitting the unitors:
\begin{center}
\input{diagrams/connector.tikz}
\end{center}

@@

\begin{proposition}
\label{prop-optic-functor}
Collectively this operation defines a \todo{pseudo?}functor \[\Optic : \SymmMonCat \to \SymmMonCat.\]
\end{proposition}
\begin{proof}
\todo{todo}
\end{proof}

@@

Categories of optics admit their own diagrammatic language, by virtue of being \emph{teleological categories}, monoidal categories with counits. In fact, we show in Theorem \ref{optic-is-free-teleological-cat} that $\Optic_\C$ is the free teleological category on a collection of dualisable morphisms $\C$.

In this section we assume that $\C$ is a \emph{strict} symmetric monoidal category. It will be convenient to equip $\Optic_\C$ with a slightly different symmetric monoidal structure: \todo{And the universal property given in this section is an argument for this being the ``morally correct'' tensor}

\begin{definition}
The \emph{switched} monoidal product on $\Optic_\C$ is given on objects by 
\begin{align*}
(S, S') \switched (T, T') := (S \otimes T, T' \otimes S')
\end{align*}
And on morphisms $\langle l, r \rangle : S \hto A$ and $\langle l', r' \rangle : T \hto B$ by:
\begin{center}
\input{diagrams/switched-tensor-on-morphisms.tikz}
\end{center}
\end{definition}

\begin{proposition}
$\Optic_\C$ equipped with the switched tensor is monoidally equivalent to it equipped with the unswitched tensor.
\end{proposition}
\begin{proof}
\todo{todo, I hope this is true?}
\end{proof}

\begin{definition}[{\cite[Definition 5.1]{CoherenceForLenses}}]
A \emph{teleological category is} a symmetric monoidal category $(\C, \otimes, I)$, equipped with:
\begin{itemize}
\item A wide symmetric monoidal subcategory $\C_d$ of \emph{dualisable morphisms}, with an involutive strong symmetric monoidal functor $(-)^* : \C_d \to \overline{\C_d^\op}$, where---not finding a standard symbol for such a thing---we mean $\overline{\C_d^\op}$ to be the category with both the direction of the arrows \emph{and} the order of the tensor flipped: $(A \otimes B)^* = B^* \otimes A^*$. \todo{Do we want this to be strict?}
\item A symmetric monoidal extranatural family of morphisms $\varepsilon_X : X \otimes X^* \to I$, called \emph{counits}, natural with respect to the morphisms in $\C_d$.
\end{itemize}
\end{definition}
Unpacking the definition, $\varepsilon$ being a symmetric monoidal extranatural transformation amounts to the following diagrams in $\C$ commuting:
\[
\begin{tikzcd}
X \otimes Y^* \ar[r, "f \otimes Y^*"]  \ar[d, "X \otimes f^*", swap] & Y \otimes Y^* \ar[d, "\varepsilon_Y"] \\
X \otimes X^* \ar[r, "\varepsilon_X", swap] & I
\end{tikzcd} \hspace{1cm}
\begin{tikzcd}
X^* \otimes X \ar[r, "\sigma"]  \ar[dr, "\varepsilon_{X^*}", swap] & X \otimes X^* \ar[d, "\varepsilon_X"] \\
& I
\end{tikzcd} \hspace{1cm}
\begin{tikzcd}
X \otimes Y \otimes Y^* \otimes X^* \ar[r, "X \otimes \varepsilon_Y \otimes X^*"]  \ar[dr, "\varepsilon_{X \otimes Y}", swap] & X \otimes X^* \ar[d, "\varepsilon_X"] \\
& I
\end{tikzcd}
\] 
where $f : X \to Y$ is in $\C_d$.

%%%% The following is a version with duality given by reflection.
%\begin{definition}[{\cite[Definition 5.1]{CoherenceForLenses}}]
%A \emph{teleological category is} a symmetric monoidal category $(\C, \otimes, I)$, equipped with:
%\begin{itemize}
%\item A wide symmetric monoidal subcategory $\C_d$ of \emph{dualisable morphisms}, with an involutive strong symmetric monoidal functor $(-)^* : \C_d \to \C_d^\op$; and,
%\item A family of morphisms $\varepsilon_X : X \otimes X^* \to I$, called \emph{counits}, natural with respect to the morphisms in $\C_d$, such that
%\[
%\begin{tikzcd}
%X \otimes Y^* \ar[r, "f \otimes Y^*"]  \ar[d, "X \otimes f^*", swap] & Y \otimes Y^* \ar[d, "\varepsilon_Y"] \\
%X \otimes X^* \ar[r, "\varepsilon_X", swap] & I
%\end{tikzcd} \hspace{1cm}
%\begin{tikzcd}
%X^* \otimes X \ar[r, "s_{X^*, X}"]  \ar[dr, "\varepsilon_{X^*}", swap] & X \otimes X^* \ar[d, "\varepsilon_X"] \\
%& I
%\end{tikzcd}
%\]
%\[
%\begin{tikzcd}
%X \otimes Y \otimes X^* \otimes Y^* \ar[r, "X \otimes s_{Y,X^*} \otimes X^*"]  \ar[dr, "\varepsilon_{X \otimes Y}", swap] & X \otimes X^* \otimes Y \otimes Y^* \ar[d, "\varepsilon_X \otimes \varepsilon_Y"] \\
%& I
%\end{tikzcd}
%\]
%commute for all $f : X \to Y$ is in $\C_d$.
%\end{itemize}
%\end{definition}

%\begin{remark}
%Compact closed categories very nearly form examples of teleological categories. In a compact closed category, there are canonical isomorphisms $(X \otimes Y)^* \cong Y^* \otimes X^*$. In a teleological category, however, duality does not reverse the order of the tensor. \todo{If $\C$ is strict, as it is here, does this matter?}
%\end{remark}

Note that because $\C_d$ is symmetric monoidal and has the same collection of objects as $\C$, the symmetric monoidal structure morphisms of $\C$ must all be in $\C_d$.

\begin{example}
Any symmetric monoidal category with terminal monoidal unit is trivially teleological, setting the dualisable morphisms to be the structure maps of the monoidal category.
\end{example}

\begin{example}
Any compact closed category is a teleological category, where every morphism is dualisable and the unit morphisms have been forgotten.
\end{example}

@@

@@

Together we have $\Tele$, the subcategory of $\SymmMonCat$ given by teleological categories and teleological functors. There are evident functors $U : \Tele \to \SymmMonCat$ and $(-)_d : \Tele \to \SymmMonCat$, that take a teleological category to its underlying symmetric monoidal category and subcategory of dualisable morphisms respectively.

\begin{proposition}
$\Optic_\C$ forms a teleological category, where:
\begin{itemize}
\item The dualisable morphisms are all morphisms of the form $\iota(f, g)$;
\item The involution is given on objects by $(S, S')^* := (S', S)$, and on a morphisms by \[\iota(f, g)^* := \iota(g, f);\]
\item The counit $\varepsilon_{(S, S')} : (S, S') \switched (S, S')^* = (S \otimes S', S \otimes S') \to (I, I)$ is given by the connector: \[\varepsilon_{(S, S')} := c_{S \otimes S'}.\]
\end{itemize}
\end{proposition}
\begin{proof}
That morphisms of the form $\iota(f, g)$ constitute a category and that $(-)^*$ is a symmetric monoidal involution is clear: we have
\begin{align*}
\left( (S, S') \switched (T, T') \right)^* 
&= \left( S \otimes T, T' \otimes S' \right)^* \\
&= \left(T' \otimes S', S \otimes T  \right) \\
&= (T', T) \otimes (S', S) \\
&= (T, T')^* \otimes (S, S')^*
\end{align*}

To check extranaturality of $\varepsilon$, suppose we have a dualisable optic $\iota(f, g) : (S, S') \hto (T, T')$, so $f : S \to T$ and $g : T' \to S'$. Extranaturality is witnessed by the equality of the string diagrams:
\begin{center}
\input{diagrams/counit-extranatural-left.tikz}
\qquad \raisebox{1.5cm}{$=$} \qquad 
\input{diagrams/counit-extranatural-right.tikz}
\end{center}
Symmetry:
\begin{center}
\includegraphics[width=.4\textwidth]{diagrams/counit-symmetry}
\end{center}
Monoidality:
\begin{center}
\includegraphics[width=.4\textwidth]{diagrams/counit-monoidal}
\end{center}

@@

%\todo{They stress about the empty set, have I avoided that issue?}

\begin{proposition}
The functor $\Optic : \SymmMonCat \to \SymmMonCat$ of Proposition \ref{prop-optic-functor} has its image in the subcategory $\Tele$.
\end{proposition}
\begin{proof}
We must show that for a symmetric monoidal functor $F : \C \to \D$, the induced functor $\Optic_\C \to \Optic_\D$ is teleological.

@@

%\begin{proposition}
%Suppose $\langle l, r \rangle : (S, S') \hto (A, A')$ has complement $M$. Then 
%\begin{align*}
%\langle l, r \rangle = (\varepsilon_{(M, I)} \otimes (A, A'))\iota(l, r).
%\end{align*}
%\end{proposition}
%\begin{proof}
%First note that because $\C$ is strict monoidal, the counit $\varepsilon_{(M, I)} : (M \otimes I, M \otimes I) = (M, M) \hto (I, I)$ is equal to the connector $c_M : (M, M) \hto (I, I)$.
%
%The claim then holds, as any optic 
%\begin{center}
%\input{diagrams/generic-optic-noline.tikz}
%\end{center}
%is equal to the composite
%\begin{center}
%\input{diagrams/optic-decomposed.tikz}
%\end{center}
%where we have drawn the normally invisible strings corresponding to the unit $I$.
%\end{proof}

\begin{proposition}
\label{prop-optic-decompose}
Suppose $\langle l, r \rangle : (S, S') \hto (A, A')$ has complement $M$. Then 
\begin{align*}
\langle l, r \rangle = (\varepsilon_{(M, I)} \otimes (A, A'))(j(l) \otimes j(r)^*)
\end{align*}
where $j : \C \to \Optic_\C$ is the functor $j(A) := \iota(A, I)$.
\end{proposition}
\begin{proof}
First note that because $\C$ is strict monoidal, the counit $\varepsilon_{(M, I)} : (M \otimes I, M \otimes I) = (M, M) \hto (I, I)$ is equal to the connector $c_M : (M, M) \hto (I, I)$.
\todo{todo}
\end{proof}

\begin{theorem}
\label{optic-is-free-teleological-cat}
Suppose $\C$ is a strict symmetric monoidal category and $\T$ is a teleological category. For all monoidal functors $F : \C \to \T_d$, there exists a teleological functor $K : \Optic_\C \to \T$, unique up to isomorphism, such that $Kj \cong F$.
%$\Optic$ is left adjoint to $(-)_d$, so $\Optic_\C$ is the free teleological category on a category of dualisable morphisms $\C$.
\end{theorem}
\begin{proof}
On objects, define $K(S, S') = FS \otimes (FS')^*$. Suppose $\langle l, r \rangle : (S, S') \hto (A, A')$ is an optic. By the previous Proposition,
\begin{align*}
\langle l, r \rangle = (\varepsilon_{(M, I)} \otimes (A, A'))(j(l) \otimes j(r)^*)
\end{align*}
But now, in order for $K$ to be a teleological functor with $Kj \cong F$, each piece of this expression must be preserved by $K$, so we are forced to define:
\begin{align*}
K\langle l, r \rangle = (\varepsilon_{FM} \otimes FA \otimes (FA')^*)(Fl \otimes (Fr)^* )
\end{align*}
%
%This is much easier to understand in the diagrammatic language for teleological categories: \todo{todo}
%
%We now have to check that this is well defined. \todo{todo}

\todo{incomplete}
\end{proof}

@@

\begin{itemize}
\item Given $f : S \to A$ and $g : A' \to S'$, the optic $j(f) \otimes j(g)^* : (S \otimes I, I \otimes S') \hto (A \otimes I, I \otimes A')$ is not equal to $\iota(f, g)$. It is also not in a form that can be composed with the unitors of $\Optic_\C$, as the $I$s are not on the same side in either the domain or codomain.
\item \todo{wasn't there more?}
\end{itemize}

@@

For a fixed $M \in \M$, there is a map $\C(S, M A) \times \C(M A, S) \to \C(S, S)$ given by composition. This induces a map $outside : \Optic_\M(S, A) \to \C(S, S)$. By composition in the other direction we have a map $\C(S, M A) \times \C(M A, S) \to \C(M A, M A)$. Including this object $\C(M A, M A)$ into the coend $\int^{M \in \M} \C(M A, M A)$, we induce a map $inside : \Optic_\M(S, A) \to \int^{M \in \M} \C(M A, M A)$. 
\begin{definition}
An optic $p : S \hto A$ is \emph{lawful} if $outside(p) = \id_S$, and $inside(p) = \langle \phi A \rangle$ for some $\phi : M \to M$ in $\M$.
\end{definition}

@@

\begin{proposition}
\label{prop-onthenose}
If $p : S \to A$ is a lawful optic, then there exists a representative $p = \langle l, r \rangle$ such that $rl = \phi A$ on the nose for some $\phi : M \to M$.
\end{proposition}
\begin{proof}
The coend relation in $\int^{M \in \M} \C(M A, M A)$ states that two morphisms $f : M A \to M A$ and $g : N A \to N A$ are identified when there exists a $k : N A \to M A$ and morphism $\phi : M \to N$ such that
\[
\begin{tikzcd}
M A \ar[r, "\phi_A"] \ar[d, "f", swap] & N A \ar[d, "g"] \ar[dl, "k"] \\
M A \ar[r, "\phi_A", swap] & N A
\end{tikzcd}
\]
commutes. This gives a binary relation $f \rightsquigarrow g$ that is not likely to be symmetric or transitive in general. The coend is obtained by identifying all morphisms related in this way, $\langle f \rangle = \langle g \rangle$ iff there exists a zigzag $f = u_1 \leftrightsquigarrow u_2 \leftrightsquigarrow \dots  \leftrightsquigarrow u_{n-1} \leftrightsquigarrow u_n = g$ where the relation may apply in either direction.

Note that if $f \rightsquigarrow g$ then $ff \rightsquigarrow gg$: this is clear by stacking the above diagram, and then taking $k' = fk = kg$. This argument can be repeated to show that $f^n \rightsquigarrow g^n$ for any $n$.

Similarly two representatives $\langle l, r \rangle : S \hto A$ and $\langle l', r' \rangle : S \hto A$ with complements $M$ and $N$ respectively are identified when there exists a $\phi : M \to N$ such that
\[
\begin{tikzcd}
M A \ar[r, "\phi_A"] \ar[d, "r", swap] & N A \ar[d, "r'"] \\
S \ar[r, equal] \ar[d, "l", swap] & S \ar[d, "l'"]  \\
MA \ar[r, "\phi_A"] & N A
\end{tikzcd}
\]
commutes.

Now suppose $p = \langle l, r \rangle : S \hto A$ is a lawful optic, so $rl = \id_S$ and $\langle lr \rangle = \langle \psi A\rangle$. There therefore must exist a finite chain of relations $lr = u_1 \leftrightsquigarrow \dots \leftrightsquigarrow u_n = \psi A$.
Supposing that the first relation faces rightwards, we have a diagram:
\[
\begin{tikzcd}
M A \ar[r, "\phi_A"] \ar[d, "r", swap] & N A \ar[dd, "u_2"] \ar[ddl, "k"] \\
S \ar[d,"l",swap] & \\
M A \ar[r, "\phi_A", swap] & N A
\end{tikzcd}
\]
Note that $r = rlr = rk\phi_A$. We therefore have the commutative diagram
\[
\begin{tikzcd}
M A \ar[r, "\phi_A"] \ar[d, "r", swap] & N A \ar[d, "rk"] \\
S \ar[r, equal] \ar[d, "l", swap] & S \ar[d, "\phi_Al"] \\
MA \ar[r, "\phi_A", swap] & N A
\end{tikzcd}
\]
Define $l' = \phi_A l$ and $r' = rk$. The above diagram is a witness that $\langle l, r \rangle = \langle l', r' \rangle$ as an optic $S \hto A$.

The composite $l'r'$ is:
\begin{align*}
l'r' &= \phi_A l r k \\
&= \phi_A k \phi_A k \\
&= u_2^2
\end{align*}

A symmetric argument shows that if instead the relation faces leftward, so $lr \leftsquigarrow g$, there exists $l'$ and $r'$ so that $\langle l, r \rangle = \langle l', r' \rangle$ and $l'r' = u_2^2$.

We can now inductively apply the above argument to the shorter chain \[l'r' = u_1^2 \leftrightsquigarrow \dots \leftrightsquigarrow u_n^2 \leftrightsquigarrow (\psi A)^2 = \psi^2 A,\] obtained by squaring each morphism in the original chain, until we are left with a representative $\langle l^*, r^* \rangle$ such that $l^*r^* = \psi^N A$, for some $N>0$. This pair $\langle l^*, r^* \rangle$ is the required representative.
\end{proof}

@@

\begin{remark}
In the case that $\C$ is cartesian closed, $\Optic_\times$ is monoidal closed via the astonishing formula
\begin{align*}
[(S, S'), (A, A')] := (\homC(S, A) \times \homC(S \times A', S'), \, S \times A')
\end{align*}
where $\homC(-, -)$ denotes the internal hom. For a proof see \cite[Section 1.2]{DialecticaCategories}. This cannot be extended to non-cartesian closed categories, the isomorphism 
\begin{align*}
\Optic_\times((S, S') \otimes (T, T'), (A, A')) \cong \Optic_\times((S, S'),  [(T, T'), (A, A')])
\end{align*}
uses the diagonal maps of $\C$ in an essential way.
\end{remark}

@@

\begin{align*}
\Optic_\otimes((S, S'), (A, A')) &= \int^{M \in \C} \C(S, M \otimes A) \times \C(M \otimes A', S') \\
&\cong \int^{M \in \C} \C(S, M \otimes A) \times \C(M, \homC(A',S')) \\
&\cong \C(S, \homC(A',S') \otimes A)
\end{align*}
where $\homC(S, A)$ denotes the internal hom. If $\C$ is cartesian, this is of course isomorphic to the pair of maps discussed in the previous section.

@@

For any category $\C$, there is a unique action of the terminal category $1$ on $\C$ that fixes every object.

\begin{proposition}
The category of optics for this action is isomorphic to $\C \times \C^\op$.
\end{proposition}
\begin{proof}
\begin{align*}
\Optic_1((S, S'), (A, A')) &= \int^{M \in 1} \C(S, MA) \times \C(MA', S') \\
&\cong \C(S, \star \cdot A) \times \C(\star \cdot A', S') \\
&\cong \C(S, A) \times \C(A', S')
\end{align*}
where $\star$ denotes the object of $1$.
\end{proof}
